I Compensate, therefore I Am \\ (accurate for DNN verification)

### 2024-04-26 lipics-master/Paper_noProof.tex
till experiments
14 pages

In this paper, we introduced the notion of {\em compensating pairs of paths}, with rationale why such a phenomenon creates inaccuracies hard to handle when verifying DNNs. We proved that this phenomenon is actually explaining entirely the inaccuracies, as in their absence, even the simplest Box abstraction (interval arithmetic) suffices to verify accurately DNNs. We verified experimentally that DNNs learnt in a natural way, harder to verify than robustly trained DNNs, exhibit paths with larger compensating strength.

Based on this idea of compensating pairs of paths, we proposed abstraction $\MILP_{Z}$ considering a subset $Z$ of the set of unstable ReLU nodes, that we proved to be fully accurate if {\em all} the compensating paths are covered by $Z$. Our empirical studies revealed that selecting $Z$ to cover only {\em a few} paths with the most significant compensation yields highly accurate results. Additionally, we observed that the runtime can be advantageous in certain scenarios, though not universally. This underscores the potential of compensating strength as an innovative and promising avenue for enhancing DNN verification.
 We finally proposed several ways to use the compensating strength to optimize current tools in different directions, and leave that as future work.

>>>>

In this paper, we introduced the notion of {\em compensating pairs of paths}, with rationale why such a phenomenon creates inaccuracies hard to handle when verifying DNNs. We proved that this phenomenon is actually explaining entirely the inaccuracies, as in their absence, even the simplest Box abstraction (interval arithmetic) suffices to verify accurately DNNs. This is experimentally confirmed by the fact that DNNs harder to verify (verification-agnostic) also exhibit paths with larger compensating strength than  robustly-trained DNNs.

Based on this idea of compensating pairs of paths, we proposed the $\MILP_{Z}$ abstraction considering a subset $Z$ of the set of unstable ReLU nodes, that we proved to be fully accurate if {\em all} the compensating paths are covered by $Z$. Our empirical studies revealed that selecting $Z$ to cover only {\em a few} paths with the most significant compensation can yield highly accurate results, up to 20\% than SOTA within the same runtime, again validating the focus on compensating strength. This underscores the potential of compensating strength as an innovative and promising avenue for enhancing DNN verification.
 We finally proposed several ways to use the compensating strength to optimize current tools in different directions, and leave that as future work.
\newpage

### 2024-04-26 lipics-master/experiments.tex
14 pages

	\caption{$\%$ of verified images and average runtime in seconds.
		Results for PRIMA, $\beta$-Crown and the upper bound on the $\%$ are from \cite{crown}.
	}
	\vspace{-0.8cm}
	\label{tab:example}
\end{table}

@@

In this work, we focused on DNNs employing ReLU activation functions since our approached on usage of MILP, which is naturally suited for ReLU activation function. However, it is worth remarking that the notion of compensation strength is independent of  activation function, and therefore, an interesting direction of future work would be to explore other the impact of compensation strength-based approaches for other activation functions.

Regarding accuracy, {\toolname} ranks highly among verifiers that abstract values. However, its runtime, particularly for larger DNNs, can be relatively slow, though not excessively so. Numerous optimization avenues exist. A primary strategy involves utilizing $\alpha$-$\beta$-CROWN initially in the sequence, reserving {\toolname} for images unresolved by $\alpha$-$\beta$-CROWN, potentially enhancing this with parallel node computation across additional CPU cores. Another tactic could involve adapting the refined branch-and-bound approach of $\beta$-CROWN to permit {\toolname} to process additional layers beyond MILP's capability, concluding the verification with BaB. More complex strategies might include modifying BaB to branch on unstable ReLU nodes, driven by compensation strength, or analyzing crucial nodes linked to uncertain outputs, enabling quicker computation of less precise bounds for less critical nodes. Furthermore, implementing a Refinement Abstraction framework, akin to methodologies in \cite{atva}, \cite{elboher}, or \cite{SRGR}, could be beneficial.

In terms of neuron correlation, a key feature explicitly encoded by PRIMA, our findings suggest that leveraging compensation strength for network abstraction is generally more effective. Although precise local accuracy is achieved by directly considering ReLU nodes, the accuracy diminishes when correlations originate from distant layers. Retaining a select few significant correlations, in the style of PRIMA, as explicit linear constraints in the MILP model, might offer a strategy to further improve accuracy.

>>>>

	\caption{$\%$ of verified images and average runtime in seconds.
		Results for PRIMA, $\beta$-Crown and the upper bound on the $\%$ are from \cite{crown}, except for $6 \times 500$ for which we run every verifier ourselves.
	}
	\vspace{-0.8cm}
	\label{tab:example}
\end{table}

@@

CMP ranks highly in terms of accuracy, on the moderatly large network (but challenging verification-agnostic) DNNs we experimented with. Even if the CMP complexity is not subject to combinatorial explosion, its runtime, particularly for larger DNNs, can be relatively slow, as CMP needs to consider every node one by one. Tackling much larger networks may thus need different implementation than CMP, while still benefitting from the novel compensation notion we uncovered and validated thanks to CMP. Numerous avenues exist:

A primary strategy involves utilizing $\alpha$-$\beta$-CROWN initially in the sequence, reserving {\toolname} for images unresolved by $\alpha$-$\beta$-CROWN, potentially enhancing this with parallel node computation across additional CPU cores. Another tactic could involve adapting the refined branch-and-bound approach of $\beta$-CROWN to permit {\toolname} to process additional layers beyond MILP's capability, concluding the verification with BaB. More complex strategies might include modifying BaB to branch on unstable ReLU nodes, driven by compensation strength, or analyzing crucial nodes linked to uncertain outputs, enabling quicker computation of less precise bounds for less critical nodes. Furthermore, implementing a Refinement Abstraction framework, akin to methodologies in \cite{atva}, \cite{elboher}, or \cite{SRGR}, could be beneficial.

In terms of neuron correlation, a key feature explicitly encoded by PRIMA, our findings suggest that leveraging compensation strength for network abstraction is generally more effective. Although precise local accuracy is achieved by directly considering ReLU nodes, the accuracy diminishes when correlations originate from distant layers. Retaining a select few significant correlations, in the style of PRIMA, as explicit linear constraints in the MILP model, might offer a strategy to further improve accuracy while keeping the number of binary variables limited.

Finally, we focused on DNNs employing ReLU activation functions since our approached on usage of MILP, which is naturally suited for ReLU activation function. However, it is worth remarking that the notion of compensation strength is independent of activation function, and therefore, an interesting direction of future work would be to explore other the impact of compensation strength-based approaches for other activation functions.
